The Talking Heads Experiment

### 2015-04-22 chap10/chap10.tex
changes to headings and minor mistakes across several chapters

\chapter{Beyond the Talking Heads Experiment}
\label{c:postscriptum}
%\newcounter{foot}
\setcounter{foot}{1}

@@

\subsubsection{Object Identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or ``anchors'' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\subsubsection{Event Identification} 
The second subsystem detects and tracks events, again at different hierarchical levels. The task is similar
to that of detecting objects but the grouping is based on changes in properties of objects rather than 
on invariances. Event detection is organised in three steps: 
\begin{enumerate}
\item Detect change by computing qualitative descriptions for the movement of an object, the contact between 
two objects, the approaching of two objects, the positioning of an object with respect to another one, etc. 
\item Detect micro-events by grouping image subsequences during which the same configuration 
of qualitative descriptors holds. 
\item Detect events which are defined as sequences of micro-events recognised using probabilistic state machines. 
For example a pick-up event (as in \figref{fig:PERACT-images}) involves three micro-events: (1) the hand 
moves towards the object, (2) the hand touches the object, and (3) both move away together. Processes 
concerned with recovering such events use a library of event definitions which is matched against the stream of micro-events.
\end{enumerate}
An example of output from this subsystem is shown below. Each event has an index (e.g. 77), a time period (e.g. 
from 780 to 803), other events or properties that have played a role in recognising this event, a 
label (e.g. {\verb Appear}), and information of the status (like ongoing or finished).

@@

\subsubsection{Qualitative Descriptors} 
The third subsystem consists of feature detectors that attempt to find qualitative descriptions for units at different levels 
of the object or event hierarchy. The result of all these processes is a set of streams, reporting objects and their 
properties dynamically in response to a changing world as well as the roles these different objects play in the events. 
A (short term) memory of these streams is kept as events unfold. It is used by the semantic system to construct or 
interpret the semantic structures that have to be expressed. Here is a small snapshot of the state of 
this visual memory with time stamps when the states occurred.

>>>>

\chapter{Beyond the Talking Heads experiment}
\label{c:postscriptum}
%\newcounter{foot}
\setcounter{foot}{1}

@@

\subsubsection{Object identification} The first subsystem attempts to detect and track visual units at different hierarchical levels. 
It detects and 'latches onto' regions in the image that are generated by objects of 
interest in the environment. This results in deictic pointers or ``anchors'' that establish and monitor indexical references 
between internal symbols and external objects. Tracking not only takes place for a single object, but for 
an open-ended set of objects at different hierarchical levels, as long as they are part of the same spatio-temporal context.

@@

\subsubsection{Event identification} 
The second subsystem detects and tracks events, again at different hierarchical levels. The task is similar
to that of detecting objects but the grouping is based on changes in properties of objects rather than 
on invariances. Event detection is organised in three steps: 
\begin{enumerate}
\item Detect change by computing qualitative descriptions for the movement of an object, the contact between 
two objects, the approaching of two objects, the positioning of an object with respect to another one, etc. 
\item Detect micro-events by grouping image subsequences during which the same configuration 
of qualitative descriptors holds. 
\item Detect events which are defined as sequences of micro-events recognised using probabilistic state machines. 
For example a pick-up event (as in \figref{fig:PERACT-images}) involves three micro-events: (1) the hand 
moves towards the object, (2) the hand touches the object, and (3) both move away together. Processes 
concerned with recovering such events use a library of event definitions which is matched against the stream of micro-events.
\end{enumerate}
An example of output from this subsystem is shown below. Each event has an index (e.g. 77), a time period (e.g. 
from 780 to 803), other events or properties that have played a role in recognising this event, a 
label (e.g. {\verb Appear}), and information of the status (like ongoing or finished).

@@

\subsubsection{Qualitative descriptors} 
The third subsystem consists of feature detectors that attempt to find qualitative descriptions for units at different levels 
of the object or event hierarchy. The result of all these processes is a set of streams, reporting objects and their 
properties dynamically in response to a changing world as well as the roles these different objects play in the events. 
A (short term) memory of these streams is kept as events unfold. It is used by the semantic system to construct or 
interpret the semantic structures that have to be expressed. Here is a small snapshot of the state of 
this visual memory with time stamps when the states occurred.

### 2015-04-22 chap8/chap8.tex
changes to headings and minor mistakes across several chapters

\chapter{The First Series (1999)}
\label{c:first-series}
%\newcounter{foot}

@@

The Laboratorium\is{Laboratorium exhibition}
exhibition was a major artistic event in Europe during the summer of 1999 organised by 
Bruno Verbergt of an organisation called 'Antwerpen Open'.
It was one of the first exhibitions that put so much emphasis on art as a research 
activity and on profound interactions between art and science.
Participants included both scientists and science historians such as Peter Galison, Bruno Latour, Israel Rosenfield
and Isabelle Stengers, as well as architects such as Rem Koolhaas and artists with an affinity for research 
such as Peter Fischli and David Weiss, Gabriel Orozco, Carsten H\"oller, Panamarenko, Lawrence Weiner a.o. 
The exhibition was visited by 8000 people.

@@

\todo{I have framed Figure 8.2 because it contains multiple images}
\begin{figure}[htbp]
 \centerline{\includegraphics[width=.35\textwidth]{chap8/figures/zoom-in-head.pdf}}
\caption{\label{fig:laboratorium}A single Talking Head at the Paris site. In the background we see (bottom) the display with an
image that the agent picked up, (on top of that) a screen displaying the processed image and the discrimination trees, 
and to the right the loudspeaker broadcasting the speech output of the agent.}
\end{figure}

A typical example of the kind of images that the camera's picked up are shown in \figref{fig:images}. 
To play a game, the agent had to find a sufficiently delineated group of objects and each object had 
to exceed a minimal size. Initially quite complex configurations were put on the white board but this made it very difficult for agents to find a coherent group and to develop good concepts for referring to the topic. The right 
image in \figref{fig:images} provides clear examples of up, down, left and right, different colors, and 
also opportunities to use multiple words (such as `red bottom').

@@

Numerous talks were given and various papers published in scientific fora. There were also various talks within 
the art context\footnote{Examples are a `gallery talk' as Salon 3 at the Elephant and Castle Centre
in London on 2 december 1998, organised by Hans Ulrich Obrist and Molly Nesbit, and a presentation together with 
Hans-Ulrich Obrist and Rem Koolhaas in Antwerp to launch the Laboratorium book on 3 October 2001.}
The experiment received wide coverage in 
the media thanks to its public exposure as part of a major exhibition. The Süddeutsche Zeitung called it 
``Angels with Internet wings''. All this lead to invitations to show the experiment
also in other venues, as discussed in the next chapter. 
Regretably, the challenge just to keep the experiment in the air with the available human resources
and the pressure for going on with new experiments prevented us from doing more adequate data gathering 
and analysis. Nevertheless some analyses appeared, particularly as carried out by Frederic Kaplan\cite{Kaplan:2001}.

>>>>

\chapter{The first series (1999)}
\label{c:first-series}
%\newcounter{foot}

@@

The Laboratorium\is{Laboratorium exhibition}
exhibition was a major artistic event in Europe during the summer of 1999 organised by 
Bruno Verbergt of an organisation called `Antwerpen Open'.
It was one of the first exhibitions that put so much emphasis on art as a research 
activity and on profound interactions between art and science.
Participants included both scientists and science historians such as Peter Galison, Bruno Latour, Israel Rosenfield
and Isabelle Stengers, as well as architects such as Rem Koolhaas and artists with an affinity for research 
such as Peter Fischli and David Weiss, Gabriel Orozco, Carsten H\"oller, Panamarenko, Lawrence Weiner a.o. 
The exhibition was visited by 8000 people.

@@

\begin{figure}[htbp]
 \centerline{\includegraphics[width=.35\textwidth]{chap8/figures/zoom-in-head.pdf}}
\caption{\label{fig:laboratorium}A single Talking Head at the Paris site. In the background we see (bottom) the display with an
image that the agent picked up, (on top of that) a screen displaying the processed image and the discrimination trees, 
and to the right the loudspeaker broadcasting the speech output of the agent.}
\end{figure}

A typical example of the kind of images that the camera's picked up are shown in \figref{fig:images}. 
To play a game, the agent had to find a sufficiently delineated group of objects and each object had \todo{I have framed Figure 8.2 because it contains multiple images}
to exceed a minimal size. Initially quite complex configurations were put on the white board but this made it very difficult for agents to find a coherent group and to develop good concepts for referring to the topic. The right 
image in \figref{fig:images} provides clear examples of up, down, left and right, different colors, and 
also opportunities to use multiple words (such as `red bottom').

@@

Numerous talks were given and various papers published in scientific fora. There were also various talks within 
the art context.\footnote{Examples are a `gallery talk' as Salon 3 at the Elephant and Castle Centre
in London on 2 December 1998, organised by Hans Ulrich Obrist and Molly Nesbit, and a presentation together with 
Hans-Ulrich Obrist and Rem Koolhaas in Antwerp to launch the Laboratorium book on 3 October 2001.}
The experiment received wide coverage in 
the media thanks to its public exposure as part of a major exhibition. The Süddeutsche Zeitung called it 
``Angels with Internet wings''. All this lead to invitations to show the experiment
also in other venues, as discussed in the next chapter. 
Regretably, the challenge just to keep the experiment in the air with the available human resources
and the pressure for going on with new experiments prevented us from doing more adequate data gathering 
and analysis. Nevertheless some analyses appeared, particularly as carried out by Frederic Kaplan (\citealt{Kaplan:2001}).

### 2015-04-22 chap9/chap9.tex
changes to headings and minor mistakes across several chapters

\chapter{The Second Series (2000--2001)}
\label{c:n01se}
%\newcounter{foot}

@@

\subsection{The Exhibition} 
\enlargethispage{2\baselineskip}
The N01SE exhibition\is{N01SE exhibition} was about information and transformation. 
Various locations in Cambridge (UK) participated: The Kettle's Yard university gallery, 
the Cambridge Whipple Museum of the History of Science, the Cambridge University Museum of 
Archaeology and Anthropology and the 
Fritzwilliam Museum. The installations ran from 22 January until 26 March 2000. There was also a site 
in London at the Wellcome Trust Two10 Gallery in Euston Road, which 
ran from 28 January to 1 May 2000. Apart from the usual press coverage for art exhibitions, articles appeared in Science\footnote{http://www.firstpulseprojects.com/sciencerev.html} and the Lancet\footnote{http://www.garnettmckeen.net/lancet.html} showing that the exhibition resonated also in the scientific press.

@@

As the N01SE exhibition was in full swing, the Palais de la d\'{e}couverte, the largest science museum in Paris, 
took the initiative to integrate the experiment as part of their running exhibition for several months. This lead 
to the design of a sophisticated framework for housing the computer equipment (see \figref{fig:palais}), 
additional educational materials explaining
what the experiment was about, and a new run in much more relaxed circumstances with therefore much more 
interesting results. This new installation started its operation during the social 
dinner for the Evolution Of Language Conference in Paris organised by Jean-Louis Desalles
on April 5, 2000. It was seen by an estimated 300,000 visitors
during its installment and an article with results of this experiment
appeared in the ``Revue du Palais de la D\'{e}couverte'' \citep{Steels:2000kaplan}.

@@

The {\itshape Mus\'ee d'Art Moderne} in Paris organised a solo exhibition of artist Olafur Eliasson entitled 
``Chaque matin je me sens diff\'erent. Chaque soir je me sens le m\^eme" between 
22 March and 12 May 2002.\cite{Scherf:2002} Olafur Eliasson is known for his thorough 
investigations of colour, such as using monochromatic light to create an artificial sun in the Modern Tate
Gallery in London (2003). Eliasson invited me to jointly work out a new interpretation of the 
Talking Heads Experiment, which became ``Look into the Box''. Nicolas Neubauer, a master student working at 
that time at the Sony Computer Science Laboratory, was the chief implementer together with Angus McIntyre.  
The set-up and results are described in detail in \cite{Steels:2004} and \cite{Neubauer:2004}.\is{Museum of Modern Art installation}

>>>>

\chapter{The second series (2000--2001)}
\label{c:n01se}
%\newcounter{foot}

@@

\subsection{The exhibition} 
\enlargethispage{2\baselineskip}
The N01SE exhibition\is{N01SE exhibition} was about information and transformation. 
Various locations in Cambridge (UK) participated: The Kettle's Yard university gallery, 
the Cambridge Whipple Museum of the History of Science, the Cambridge University Museum of 
Archaeology and Anthropology and the 
Fritzwilliam Museum. The installations ran from 22 January until 26 March 2000. There was also a site 
in London at the Wellcome Trust Two10 Gallery in Euston Road, which 
ran from 28 January to 1 May 2000. Apart from the usual press coverage for art exhibitions, articles appeared in Science\footnote{http://www.firstpulseprojects.com/sciencerev.html} and the Lancet\footnote{http://www.garnettmckeen.net/lancet.html} showing that the exhibition resonated also in the scientific press.

@@

As the N01SE exhibition was in full swing, the Palais de la d\'{e}couverte, the largest science museum in Paris, 
took the initiative to integrate the experiment as part of their running exhibition for several months. This lead 
to the design of a sophisticated framework for housing the computer equipment (see \figref{fig:palais}), 
additional educational materials explaining
what the experiment was about, and a new run in much more relaxed circumstances with therefore much more 
interesting results. This new installation started its operation during the social 
dinner for the Evolution Of Language Conference in Paris organised by Jean-Louis Desalles
on 5 April 2000. It was seen by an estimated 300,000 visitors
during its installment and an article with results of this experiment
appeared in the ``Revue du Palais de la D\'{e}couverte'' \citep{Steels:2000kaplan}.

@@

The {\itshape Mus\'ee d'Art Moderne} in Paris organised a solo exhibition of artist Olafur Eliasson entitled 
``Chaque matin je me sens diff\'erent. Chaque soir je me sens le m\^eme" between 
22 March and 12 May 2002. (\citealt{Scherf:2002}) Olafur Eliasson is known for his thorough 
investigations of colour, such as using monochromatic light to create an artificial sun in the Modern Tate
Gallery in London (2003). Eliasson invited me to jointly work out a new interpretation of the 
Talking Heads Experiment, which became ``Look into the Box''. Nicolas Neubauer, a master student working at 
that time at the Sony Computer Science Laboratory, was the chief implementer together with Angus McIntyre.  
The set-up and results are described in detail in \cite{Steels:2004} and \cite{Neubauer:2004}.\is{Museum of Modern Art installation}
